---
title: "Gods_estimator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gods_estimator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = F}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")#, dev = "pdf")
library(MeanRarity)
library(patchwork) # amazing ggplot2 add-on for multi-panel plots
library(dplyr) # part of the tidyverse
library(ggplot2) # part of the tidyverse
library(tidyr) # part of the tidyverse
library(purrr) # part of the tidyverse
library(tictoc) # lightweight package for convenient timing
library(furrr) # `tidyverse` and `parallel` combined
```
Hill diversities (also called "Hill numbers") have become standard for measuring
diversity in ecology (cite the Ellison 2010 special issue). While many
interpretations of Hill numbers have been discussed, including their
relationship to species richness, Shannon’s entropy, and Simpson’s
concentration, The package `MeanRarity` is designed to illustrate an alternative
interpretation of Hill numbers as means of species rarity (the reciprocal of
relative abundance) in a community. In this interpretation, the control
parameter $\ell$ moves the index from arithmetic to geometric to harmonic means.
Thus, Hill diversites differ in the scale on which species rarity is considered,
but always weights species the same way (by their abundance) regardless of the
scale being used. This interpretation is distinct from the common idea that the
Hill parameter controls how rare vs. common species are “weighted.”

In this vignette, we show that the mean rarity perspective suggests a novel
approach to estimating true diversity from a sample, using the fact that the
sample mean is an unbiased estimate of the population mean. Although how to
estimate a species’ rarity from sample data remains unsolved, we test the
promise of this idea using what we call God’s estimator: All-knowing, God can
plug in the true population parameters for each species. Nevertheless, maybe for
fun, God chooses to compute the sample mean, weighting species’ true rarities by
their sampled frequency.

Let's see how God's estimator works!

First, we simulate some SADs with known diversity using the function `fit_SAD()`

```{r simulate species abundance distributions}

# simulate SADs for analysis
SADs_list<-map(c("lnorm", "gamma"), function(distr){
  map(c(100, 200), function(rich){
    map(c(.1,.15,.25,.5,0.75,.85), function(simp_Prop){ #simp_Prop is evenness
      fit_SAD(rich = rich, simpson = simp_Prop*(rich-1)+1, distr = distr)
    })
  })
})
# fiddled with this because of bizarre behavior. I think there is R precision
# issues if the species are too rare this occurred with some of the gamma
# distributions. I first of all set the lease even SAD to be more even, and also
# decided to drop SADs 19-20 after taking a quick look at those, which had
# rarities below 1e-9
```

Now, let's test estimator performance.We'll compare what we've called "God's 
estimator" to the naïve estimator (the mean of the observed rarities) and to a
non-parametric, asymptotic diversity estimator (CITE Chao and Jost 2015). To 
assess performance, let's look at the root mean squared error, and also bias.

```{r test estimators}

nreps <- 999

ss_to_test <- floor(10^seq(1,3, 0.2))

flatsads <- flatten(flatten(SADs_list))
nc <- parallel::detectCores() - 1

future::plan(strategy = "multiprocess", workers = nc)
options <- furrr::furrr_options(seed = TRUE)
tic()
compare_ests <- map_dfr(1:24, function(SAD){
  map_dfr(-1:1, function(ell){
   truth = as.numeric(flatsads[[SAD]][[2]][2-ell])
   truep = flatsads[[SAD]][[3]]
   dinfo = data.frame(t(flatsads[[SAD]][[1]]))
  future_map_dfr(.x=1:nreps, .f=function(nrep){
      map_dfr(ss_to_test, function(ss){
        subsam = sample_infinite(truep, ss)
        chaoest = Chao_Hill_abu(subsam, ell)
        naive = rarity(subsam, ell)
        gods = GUE(subsam, truep, ell)
        return(bind_cols(data.frame(truth, chaoest, naive, gods, n=ss, ell, SAD), dinfo))
      })
    }
    , .options= options)
  })
})
toc() # 6 mins on 2.9 GHz i7 quad core processor

# define a function for computing the root mean square
nasum <- function(x){sum(x, na.rm =T)}
rootms <- function(x){sqrt(nasum(((x)^2)/length(x)))}
namean <-function(x){mean(x, na.rm =T)}

errs <- compare_ests %>% 
  group_by(ell, distribution, fitted.parameter, n, SAD) %>% 
  mutate(godsError = gods - truth
         , naiveError = naive - truth
         , chaoError = chaoest - truth) %>% 
  summarize_at(.vars = c("godsError", "naiveError", "chaoError"), .funs = c(rootms, namean)) %>% 
  pivot_longer(godsError_fn1:chaoError_fn2,
    names_to = c("estimator", ".value"),
    names_sep = "_"
      ) %>% 
  rename(rmse = fn1, bias = fn2)
 
               
```

```{r plot rmse}
# data to make plot prettier
inds <- data.frame("ell" = c(1, 0, -1), divind = factor(c("richness"
                                                           , "Hill-Shannon"
                                                           , "Hill-Simpson")
                                                 , levels = c("richness"
                                                              , "Hill-Shannon"
                                                            , "Hill-Simpson")))
#plot
errs %>% 
  mutate(estimator = gsub("*Error", "", estimator)) %>%  
  left_join(inds) %>% 
  filter(SAD %in% c("8","10","12","19","22", "24")) %>%
  ggplot(aes(n, rmse, color = estimator))+
  geom_line(alpha = 0.7, size =1.5) +
  facet_grid(divind~SAD, scales = "free") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "individuals", y = "root mean squared error")
  
```
```{r plot bias}


errs %>% 
  mutate(estimator = gsub("*Error", "", estimator)) %>%  
  left_join(inds) %>% 
  filter(SAD %in% c("7","10","12","21","23","24")) %>%
  ggplot(aes(n, bias, color = estimator)) +
  geom_line(alpha = 0.7, size =1.5) +
  facet_grid(divind~SAD, scales = "free") +
  geom_hline(yintercept = 0) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "individuals")
  
```


In our simulations, God’s estimator was both less biased (in fact, it is
unbiased) and more accurate, across variation in SAD shape,
sample sizes, and values of the Hill control parameter. This is expected:
God’s estimator uses information not available to other methods (nor to mortal
ecologists). However, these results suggest that improved heuristic estimates of
species rarities might enable enable better diversity estimates, and pave a
route to more principled confidence intervals for Hill numbers.
