---
title: "Gods estimator of mean rarity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gods_estimator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = F}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")#, dev = "pdf")
library(MeanRarity)
library(patchwork) # amazing ggplot2 add-on for multi-panel plots
library(dplyr) # part of the tidyverse
library(ggplot2) # part of the tidyverse
library(tidyr) # part of the tidyverse
library(purrr) # part of the tidyverse
#library(tictoc) # lightweight package for convenient timing
library(furrr) # `tidyverse` and `parallel` combined
```
Hill diversities (also called "Hill numbers") have become standard for measuring
diversity in ecology (cite the Ellison 2010 special issue). While many
interpretations of Hill numbers have been discussed, including their
relationship to species richness, Shannon’s entropy, and Simpson’s
concentration, The package `MeanRarity` is designed to illustrate an alternative
interpretation of Hill numbers as means of species rarity (the reciprocal of
relative abundance) in a community. In this interpretation, the control
parameter $\ell$ moves the index from arithmetic to geometric to harmonic means.
Thus, Hill diversities differ in the scale on which species rarity is considered,
but always weights species the same way (by their abundance) regardless of the
scale being used. This interpretation is distinct from the common idea that the
Hill parameter controls how rare vs. common species are “weighted.”

In this vignette, we show that the mean rarity perspective suggests a novel
approach to estimating true diversity from a sample, using the fact that the
sample mean is an unbiased estimate of the population mean. Although how to
estimate a species’ rarity from sample data remains unsolved, we test the
promise of this idea using what we call God’s estimator: All-knowing, God can
plug in the true population parameters for each species. Nevertheless, maybe for
fun, God chooses to compute the sample mean, weighting species’ true rarities by
their sampled frequency.

Let's see how God's estimator works!

First, we simulate some SADs with known diversity using the function `fit_SAD()`

```{r simulate species abundance distributions}

# simulate SADs for analysis
SADs_list<-map(c("lnorm", "gamma"), function(distr){
  map(c(100, 200), function(rich){
    map(c(.1,.15,.25,.5,0.75,.85), function(simp_Prop){ #simp_Prop is evenness
      fit_SAD(rich = rich, simpson = simp_Prop*(rich-1)+1, distr = distr)
    })
  })
})

```

Now, let's test estimator performance.We'll compare what we've called "God's 
estimator" to the naïve estimator (the mean of the observed rarities) and to a
non-parametric, asymptotic diversity estimator (CITE Chao and Jost 2015). To 
assess performance, let's look at the root mean squared error, and also bias.

We won't run the first part as it could take a while.
```{r test estimators, eval = F}

nreps <- 9999

ss_to_test <- floor(10^seq(1,3, 0.2))

flatsads <- flatten(flatten(SADs_list))

nc <- parallel::detectCores() - 1
future::plan(strategy = "multiprocess", workers = nc)
# options <- furrr::furrr_options(seed = TRUE)
tic()
compare_ests <- map_dfr(1:24, function(SAD){
  map_dfr(-1:1, function(ell){
   truth = as.numeric(flatsads[[SAD]][[2]][2-ell])
   truep = flatsads[[SAD]][[3]]
   dinfo = data.frame(t(flatsads[[SAD]][[1]]))
  future_map_dfr(.x=1:nreps, .f=function(nrep){
      map_dfr(ss_to_test, function(ss){
        subsam = sample_infinite(truep, ss)
        chaoest = Chao_Hill_abu(subsam, ell)
        naive = rarity(subsam, ell)
        gods = GUE(subsam, truep, ell)
        return(bind_cols(data.frame(truth, chaoest, naive, gods, n=ss, ell, SAD), dinfo))
      })
    }
    # , .options= options
    )
  })
})
toc() # 6 mins on 2.9 GHz i7 quad core processor
# define a function for computing the root mean square
nasum <- function(x){sum(x, na.rm =T)}
rootms <- function(x){sqrt(nasum(((x)^2)/length(x)))}
namean <- function(x){mean(x, na.rm =T)}

errs <- compare_ests %>% 
  group_by(ell, distribution, fitted.parameter, n, SAD) %>% 
  mutate(godsError = gods - truth
         , naiveError = naive - truth
         , chaoError = chaoest - truth) %>% 
  summarize_at(.vars = c("godsError", "naiveError", "chaoError"), .funs = c(rootms, namean)) %>% 
  pivot_longer(godsError_fn1:chaoError_fn2,
    names_to = c("estimator", ".value"),
    names_sep = "_"
      ) %>% 
  rename(rmse = fn1, bias = fn2)
```

```{r plot rmse, fig.height =5, fig.width =7}
# data to make plot prettier
inds <- data.frame("ell" = c(1, 0, -1), divind = factor(c("richness"
                                                           , "Hill-Shannon"
                                                           , "Hill-Simpson")
                                                 , levels = c("richness"
                                                              , "Hill-Shannon"
                                                            , "Hill-Simpson")))
#plot
errs %>% 
  mutate(estimator = gsub("*Error", "", estimator)) %>%  
  left_join(inds) %>% 
  filter(SAD %in% c("8","10","12","19","22", "24")) %>%
  ggplot(aes(n, rmse, color = estimator))+
  geom_line(alpha = 0.7, size =1.5) +
  facet_grid(divind~SAD, scales = "free") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "individuals", y = "root mean squared error")
  
```
```{r plot bias, fig.height =5, fig.width =7}


errs %>% 
  mutate(estimator = gsub("*Error", "", estimator)) %>%  
  left_join(inds) %>% 
  filter(SAD %in% c("7","10","12","19","22","24")) %>%
  ggplot(aes(n, bias, color = estimator)) +
  geom_line(alpha = 0.7, size =1.5) +
  facet_grid(divind~SAD, scales = "free") +
  geom_hline(yintercept = 0) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "individuals")
  
```


In our simulations, God’s estimator was usually both less biased (in fact, it is
unbiased) and more accurate, across variation in SAD shape,
sample sizes, and values of the Hill control parameter. This is expected:
God’s estimator uses information not available to other methods (nor to mortal
ecologists). 

For SAD 19 we see very large RMSE for species richness. This is because God, but
not the other estimators, can distinguish the merely rare from the exceedingly
rare, and SAD 19 has wildly rare species (a quarter of species have relative
abundance of < 1/ 10 million). When these species are included randomly in a
sample, the mean rarity of that sample is very high. Our simulation may be
untroubling then, even for proponents of God's estimator. When ecologists sample
SADs with hundreds of millions or billions of individuals (e.g. Rhizospheres),
they likely do not attempt to quantify richness, given that it cannot be
estimated reliably (CITE Haegemen et al. 2013),  and also obtain reads from many
more than a few hundred individuals. For God's estimator not to be unduly
swayed by the extraordinarily rare species, however, sample sizes will have to
be extremely large. In the presence of such extreme rarity, God's estimator
remains unbiased, but will likely tend towards greater error than estimators
that do not account for rarity orders of magnitude higher than the observed
rarity in a sample. 

```{r Gods estimator in the face of extreme rarity, eval =F}
#we can check out whether upping sample sizes by a factor of 100 is 
# sufficient to bring God's estimator in line, but it's pretty time consuming to 
# do this sampling, and with the extreme rarities in SAD 19 we don't actually 
# see God's estimator behaving well with sample sizes ≤1e7, so we won't do this 
# with brute force.
# nreps <- 999
# 
# ss_to_test <- floor(10^seq(5,7, 0.25))
# SAD<-19
# 
# 
# nc <- parallel::detectCores() - 1
# future::plan(strategy = "multiprocess", workers = nc)
# # options <- furrr::furrr_options(seed = TRUE)
# tic()
# God_with_microbes <- map_dfr(-1:1, function(ell){
#    truth = as.numeric(flatsads[[SAD]][[2]][2-ell])
#    truep = flatsads[[SAD]][[3]]
#    dinfo = data.frame(t(flatsads[[SAD]][[1]]))
#   future_map_dfr(.x=1:nreps, .f=function(nrep){
#       map_dfr(ss_to_test, function(ss){
#         subsam = sample_infinite(truep, ss)
#         chaoest = Chao_Hill_abu(subsam, ell)
#         naive = rarity(subsam, ell)
#         gods = GUE(subsam, truep, ell)
#         return(bind_cols(data.frame(truth, chaoest, naive, gods, n=ss, ell, SAD), dinfo))
#       })
#     }
#     # , .options= options
#     )
#  })
# toc()
# 
# microbe_errs <- God_with_microbes %>% 
#   group_by(ell, distribution, fitted.parameter, n, SAD) %>% 
#   mutate(godsError = gods - truth
#          , naiveError = naive - truth
#          , chaoError = chaoest - truth) %>% 
#   summarize_at(.vars = c("godsError", "naiveError", "chaoError"), .funs = c(rootms, namean)) %>% 
#   pivot_longer(godsError_fn1:chaoError_fn2,
#     names_to = c("estimator", ".value"),
#     names_sep = "_"
#       ) %>% 
#   rename(rmse = fn1, bias = fn2)
# 
# 
# microbe_errs %>% 
#   mutate(estimator = gsub("*Error", "", estimator)) %>%  
#   left_join(inds) %>% 
#   ggplot(aes(n, rmse, color = estimator)) +
#   geom_line(alpha = 0.7, size =1.5) +
#   facet_grid(~divind, scales = "free") +
#   geom_hline(yintercept = 0) +
#   theme_classic() +
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(x = "individuals")
# 
# microbe_errs %>% 
#   mutate(estimator = gsub("*Error", "", estimator)) %>%  
#   left_join(inds) %>% 
#   ggplot(aes(n, bias, color = estimator)) +
#   geom_line(alpha = 0.7, size =1.5) +
#   facet_grid(~divind, scales = "free") +
#   geom_hline(yintercept = 0) +
#   theme_classic() +
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(x = "individuals")
````



What are the implications of this for mortals? These simulations suggest that
developing heuristics to estimate true species rarities might enable 
better diversity estimates, and pave a route to more principled confidence
intervals for Hill numbers than is likely possible with the non-parametric
estimators more commonly used today. While it is hard to imagine how an
estimator for true rarity could accurately infer rarities far below n/1 (where n
is sample size), our simulations (and logic) suggest that arithmetic mean rarity
(i.e. richness), highly sensitive to the stochastic inclusion of these in
samples, at bests presents a bias/variance trade-off between a mean rarity
estimator (like God's, but with estimated rather than true rarities) and more 
traditional asymptotic diversity estimators.
