---
title: "Using_MeanRarity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using_MeanRarity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r setup, include =F}
library(MeanRarity)
```
# Purpose

This R package allows estimating and visualizing Hill diversity (aka "Hill
numbers") in terms of the average species "rarity." Rarity is defined as the 
reciprocal of relative abundance.

The average species rarity may be simpler to interpret than a transformed,
generalized entropy (this is how Hill numbers are typically derived). The rarity
parameterization clarifies the role of the scaling exponent, and this R package
provides illustration tools to visualize that scaling concretely. Furthermore,
this framework may provide an entry point for developing and testing diversity
estimators and their related interval estimates (e.g., confidence intervals);
this package contains tools to simulate species abundance distributions that
should streamline that process.

# Computing Mean Rarity
`rarity()` computes Hill diversity of a sample, parameterized in terms of the
mean rarity. 

We define Hill diversity $D$ as the mean species rarity in the community

$$D=\left( \sum_{i=1}^{S}p_{i}(r_{i})^{\ell} \right)^{1/\ell} \tag{1}$$

where $D$ is diversity or mean rarity, $p_{i}$ is the relative abundance of species $i$, $r_{i}$ is the rarity of species $i$ (defined as the reciprocal of $p_{i}$), $S$ is the total species richness, and $\ell$ is the scaling exponent that determines the type of mean computed (Roswell et al. 2021). We define the expression for  $\ell = 0$  based on the limit (from left and right) as (DONT FORGET TO ADD EQN 3).

Hill diversity is more commonly written as 

$$D=\left( \sum_{i=1}^{S}p_{i}^{q} \right)^{1/1-q} \tag{2}$$ 

These expressions are equivalent when $\ell = 1-q$. Re-parameterizing Hill
diversity in terms of species rarity can provide conceptual clarity. Equation 1
is a weighted mean rarity, where the weights (relative abundances, $p_i$) are
factored out. Simplifying (as in Equation 2) hides the weighting by relative
abundance, and leads to the claim that the scaling exponent is the "weight." The
concrete meaning of "weight" when applied to $q$ in equation 2 may be difficult
to understand.

Because Equations 1 and 2 are algebraically equivalent, `rarity(x, l)` is
equivalent to `vegan::renyi(x, scale = 1-l, hill = T)` and the `qD` output from `iNEXT::estimateD(x)` where `order = 1-l`.



```{r compute rarity}
# for a totally even distribution, Hill diversity is the same regardless of `l`

even_comm <- rep(2, 6)
rarity(even_comm, 1) # richness
rarity(even_comm, -1) # Hill-Simpson

all.equal(rarity(even_comm, 0), rarity(even_comm, 0.34), rarity(even_comm, 1))

# for uneven communities, Hill diversity depends on `l`

uneven_comm <- c(20, 8, 5, 4, 2, 1)

l_vals <- c(1, 0, -1, 0.34)


data.frame(l = l_vals
           , Hill = sapply(l_vals, function(l){
                    rarity(uneven_comm, l)})
)

```



# Interpreting Mean Rarity

One way of interpreting Hill diversities is that they express the diversity of a community in terms of an imaginary community with that same diversity, but in which all species are equally abundant (Jost 2006). In this perspective, the community from `uneven_comm` is as diverse as a community of six equally abundant species (richness, `l = 1`), or four equally abundant species (Hill-Shannon diversity, `l = 0`), or three (Hill-Simpson diversity, `l = -1`). Obviously, the equivalency depends on the distribution of relative abundances, and the parameter `l`. So when `l = 1`, this community is as diverse (or as species-rich) as any community with six species. When `l = -1`, it is as diverse as a community with twice the number of species and tremendous unevenness, or half the number of species and perfect evenness:

``` {r effective numbers}
uneven_12 <- c(531, 186, 101,  63, 41, 28, 19, 13, 9, 5, 3, 1)
even_3 <- c(20, 20, 20)

rarity(uneven_comm, -1)
rarity(uneven_12, -1)
rarity(even_3, -1)

```


One might wonder what the meaning of $\ell$ is, in the effective numbers framework. Many authors have described $\ell$ as a factor that weights species relative abundances. Another interpretation (developed in this package) is that $\ell$ **rescales** species *rarities*. This interpretation lends itself to concrete visualizations of Hill diversities as mean rarities, provided by `rarity_plot()`.  

`rarity_plot()` makes balance plots to visualize mean rarity. The idea behind a balance plot is that means measure the weighted center, or a balance point, of a set of values. Of course, if you have a plank with many weights on it, where the balance point lies depends not only on the weights, but also on their relative locations... the farther towards the ends of the plank a weight lies, the more leverage it has. When describing diversity, there are simple rules for the species weights ([relative] abundances) and where they lie (their *scaled* rarities). The rarity scaling is given by the exponent, `l`.  

In a balance plot, the species abundances are plotted on the y-axis, species rarities on the x-axis (values are back-transformed, but scaling is dictated by `l`), the mean is shown with a triangular "fulcrum," and reference points are illustrated for the three Pythagorean means (i.e., arithmetic, geometric, and harmonic, given when `l= c(1, 0, -1)`, and equal to richness, Hill-Shannon diversity, and Hill-Simpson diversity). By default, each individual is plotted as a "block", and species with the same rarity are stacked vertically. 

```{r Balance plot, fig.width = 3, fig.height = 3}

rich_bal <- rarity_plot(ab = uneven_comm, l = 1, base_size = 12) +
  ggplot2::labs(x = "") +
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2") 

shan_bal <- white_y(rarity_plot(ab = uneven_comm, l = 0, base_size = 12, verbose = F)) + #do not return all the text this time
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2")

simp_bal <- white_y(rarity_plot(ab = uneven_comm, l = -1, base_size = 12, verbose = F)) + 
  ggplot2::labs(x = "") +
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2")

rich_bal 
shan_bal
simp_bal

```

# Simulating communities with known diversity

`MeanRarity::fit_SAD()` simulates a species abundance distribution with known
richness, Hill-Simpson diversity, and a distributional assumption (lognormal or
gamma). It returns parameters of the chosen probability distribution, the
richness, Hill-Shannon, and Hill-Simpson diversity of the simulated species
abundance distribution, and relative abundances for each simulated species. It
provides a warning when relative abundances are very low, implying large total
community abundance.

Unfortunately, it is difficult to assess the performance of diversity estimators
with real-world data, because ecological communities are not typically fully
censused, and therefore the true diversity of those communities is not known.
There are, of course, a few exception datasets... mostly observations of
vascular plants, often trees, where the species identity and abundances in each
plot is known (CITE tropical tree dataset). However, these may not be
representative of the variability in species abundance distributions found in
nature. To enable a more comprehensive view, `MeanRarity` includes a tool to
simulate species abundance distributions where diversity values are given as
parameters. This way, diversity (in this case, richness and Hill-Simpson
diversity) are known and fixed, rather than epiphenomena that fall out of the
simulations. In order to achieve this, we simulate species abundance
distributions with known, finite true diversity, but infinite
abundance, from which finite samples may be taken. The diversity of a given
sample from such as SAD is, of course, not pre-determined.

This approach differs from species abundance distribution simulations
implemented in other R packages. For example, `mobsim` (CITE) and `sads` (CITE)
allow users to specify continuous parametric probability distributions, but then
explicitly take finite, discrete samples from those distributions in the
simulation (in `mobsim` these can even be spatially explicit!). These are
powerful simulation tools, but one drawback from using them is that the
diversity resulting from any given set of parameter values is variable.

`fit_SAD()`, by contrast, does not simulate individuals at all. Instead, it
always gives precisely the same relative abundances for a given set of
parameters, which may be a helpful tool for exploring the behavior of diversity
metrics under different community size and sampling scenarios. 

```{r fit species abundance distributions}
# set target richness and Hill-Simpson diversity levels
rich <- 20
simpson <- 12
even_SAD <- fit_SAD(rich = rich, simpson = simpson, distr = "lnorm")

#output is a list with three elements
even_SAD

rich <- 20
simpson <- 6
uneven_SAD <- fit_SAD(rich = rich, simpson = simpson, distr = "lnorm")



ev<-radplot(even_SAD$rel_abundances) + 
  ggplot2::labs(y = "relative abundance", title =NULL)+
  ggplot2::ylim(c(0, 0.4))
un<-radplot(uneven_SAD$rel_abundances) + 
  ggplot2::labs(ylab = "relative abundance", title =NULL) +
  ggplot2::ylim(c(0, 0.4))

ev
un
```

`fit_SAD()` generates an ideal species abundance distribution, with relative
abundances for a community with infinite abundance. To take a finite sample, we
use `sample_infinite()` (so named because it samples an infinite SAD;
`sample_finite()` is for sub-sampling from a realized community/finite sample).
The realized samples have Poisson-type noise about the mean species relative
abundances given by `fit_SAD()` The diversity of a sample asymptotically
approaches the simulated diversity as sample size approaches infinity.

```{r sample from SAD}

# diversity of a small sample unlikely to match the simulated diversity values

# first, sample from our simulated SADs

even_sample <- sample_infinite(even_SAD$rel_abundances, 50)
uneven_sample <- sample_infinite(uneven_SAD$rel_abundances, 50)
#sample_infinite returns a numeric vector of sample species abundances
even_sample
uneven_sample

# compare diversity from sample vs. ideal SAD: 
# likely pretty far off for a small sample!

# with an even SAD, all diversities lower
even_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(even_sample, l)
}))

# with an uneven SAD, sample Hill-Simpson can be higher than true Hill-Simpson 
# richness is always lower
uneven_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(uneven_sample, l)
}))


big_even_sample <- sample_infinite(even_SAD$rel_abundances, 1e7)
big_uneven_sample <- sample_infinite(uneven_SAD$rel_abundances, 1e7)

# with a big enough sample, simulated diversities nearly recovered
even_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(big_even_sample, l)
}))

# with an uneven SAD, sample Hill-Simpson can be higher than true Hill-Simpson 
# richness is always lower
uneven_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(big_uneven_sample, l)
}))

```



