---
title: "Using MeanRarity"
output: rmarkdown::html_vignette
bibliography: '`r system.file( "REFERENCES.bib", package="MeanRarity")`'
vignette: >
  %\VignetteIndexEntry{Using MeanRarity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


<!-- need to explore output formats for size, speed, and features (e.g. eqn numbers) -->
<!-- had been rmarkdown::html_vignette -->
<!-- also see html_document -->
<!-- also see prettydoc::html_pretty which might solve the equation numbering issues -->
<!-- https://bookdown.org/yihui/rmarkdown/prettydoc.html -->



\renewcommand{\[}{\begin{equation}}
\renewcommand{\]}{\end{equation}}

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r setup, include =F}
library(MeanRarity)
```

# INTRODUCTION

Biodiversity underpins human existence and drives the character and function of
earth’s ecosystems. At the same time, biodiversity is shifting and arguably,
declining, at an unprecedented rate due to anthropogenic global change. Yet,
detecting and describing this change means choosing just what to measure, and
how. Ecologists are faced with a dazzling array of choices.

There is an increasing consensus that Hill diversity is the preferred way to
measure not only the species diversity of a community [@Roswell2020],
differentiation among communities [@Jost2007a, @Ellison2010, @Chao2016,
@Botta-Dukát2018, @Chao2019b, @Ohlmann2019], functional and phylogenetic
diversity [@Chao2014b, @Kang2016], genetic diversity [@Jost2008, @Sherwin2017,
@Alberdi2019], and evenness [@Chao2019a]. Although Hill diversity ties together
disparate quantatitve and disciplinary threads, unifying versions of the
most-used and -studied diversity metrics with a single equation with a single
control parameter, its adoption within ecological literature has been somewhat
slow, perhaps because visualizing and interpreting Hill diversity and the
meaning of the control parameter is difficult.

Here, we introduce the R package `MeanRarity`, which enables estimating and
visualizing Hill diversity (aka "Hill numbers") in terms of the average species
"rarity." Rarity is defined as the reciprocal of relative abundance.

The average species rarity may be simpler to interpret than a transformed,
generalized entropy (i.e., how Hill numbers are typically derived). The rarity
parameterization clarifies the role of the scaling exponent, and this R package
provides illustration tools to visualize that scaling concretely. Furthermore,
this framework may provide an entry point for developing and testing diversity
estimators and their related interval estimates (e.g., confidence intervals);
this package contains tools to simulate species abundance distributions that
should streamline that process.

This R package provides functions for three purposes: (1) computing Hill
diversity as "Mean Rarity", (2) visualizing and interpreting "Mean Rarity", and
(3) simulating species abundance distributions with known diversity. The package
is availabe online on Github (see Data accessibility). An interactive shiny
application for building "balance plots" is also available online
(https://mean-rarity.shinyapps.io/rshiny_app1/).

# PACKAGE DESCRIPTION
## Computing Mean Rarity
`rarity()` computes Hill diversity of a sample, parameterized in terms of the
mean rarity. 

We define Hill diversity $D$ as the mean species rarity in the community

$$D=\left( \sum_{i=1}^{S}p_{i}(r_{i})^{\ell} \right)^{1/\ell} \tag{1}$$

where $D$ is diversity or mean rarity, $p_{i}$ is the relative abundance of
species $i$, $r_{i}$ is the rarity of species $i$ (defined as the reciprocal of
$p_{i}$), $S$ is the total species richness, and $\ell$ is the scaling exponent
that determines the type of mean computed [@Roswell2020]. We define the
expression for  $\ell = 0$  based on the limit (from left and right) as

$$D=e^{ \sum_{i=1}^{S}p_{i} * \log(r_{i})} \tag{2}$$

Hill diversity is more commonly written as 

$$D=\left( \sum_{i=1}^{S}p_{i}^{q} \right)^{1/1-q} \tag{3}$$ 

Defined when q = 1 as
$$D=e^{ -\sum_{i=1}^{S} p_{i} * \log(p_{i})} \tag{4}$$
Equations 1 and 3 (or 2 and 4) are equivalent when $\ell = 1-q$.
Re-parameterizing Hill diversity in terms of species rarity can provide
conceptual clarity. Equation 1 is a weighted mean rarity, where the weights
(relative abundances, $p_i$) are factored out. Simplifying (as in Equation 3)
hides the weighting by relative abundance, and leads to the claim that the
scaling exponent is the "weight." The concrete meaning of "weight" when applied
to $q$ in equation 2 may be difficult to understand.

Because Equations 1 and 2 are algebraically equivalent, `rarity(x, l)` is
equivalent to `vegan::renyi(x, scale = 1-l, hill = T)` and the `qD` output from `iNEXT::estimateD(x)` where `order = 1-l`.



```{r compute rarity}
# for a totally even distribution, Hill diversity is the same regardless of `l`

even_comm <- rep(2, 6)
rarity(even_comm, 1) # richness
rarity(even_comm, -1) # Hill-Simpson

all.equal(rarity(even_comm, 0), rarity(even_comm, 0.34), rarity(even_comm, 1))

# for uneven communities, Hill diversity depends on `l`

uneven_comm <- c(20, 8, 5, 4, 2, 1)

l_vals <- c(1, 0, -1, 0.34)


data.frame(l = l_vals
           , Hill = sapply(l_vals, function(l){
                    rarity(uneven_comm, l)})
)

```

## Interpreting Mean Rarity

One way of interpreting Hill diversities is that they express the diversity of a
community in terms of an imaginary community with that same diversity, but in
which all species are equally abundant [@Jost2006]. In this perspective, the
uneven community `uneven_comm` is as diverse as a community of six equally
abundant species (richness, `l = 1`), or four equally abundant species
(Hill-Shannon diversity, `l = 0`), or three (Hill-Simpson diversity, `l = -1`).
These equivalencies depend on the the distribution of relative abundances and
the parameter `l`. So when `l = 1`, this community is as diverse (or as
species-rich) as any community with six species. When `l = -1`, it is about as diverse
as a community with twice the number of species and tremendous unevenness, or
half the number of species and perfect evenness:

``` {r effective numbers}
uneven_12 <- c(531, 186, 101,  63, 41, 28, 19, 13, 9, 5, 3, 1)
even_3 <- c(20, 20, 20)

rarity(uneven_comm, -1) %>% round(2)
rarity(uneven_12, -1) %>% round(2)
rarity(even_3, -1) %>% round(2)

```


One might wonder what the meaning of $\ell$ is, in the effective numbers
framework. Many authors have described $\ell$ as a factor that weights species
relative abundances. This suggests that downweighting rare species means a lower
effective number of species, but upweighting them means a higher effective
number of species. Another interpretation (developed in this package) is that
$\ell$ **rescales** species *rarities*. This interpretation lends itself to
concrete visualizations of Hill diversities as mean rarities, provided by
`rarity_plot()`.

`rarity_plot()` makes balance plots to visualize mean rarity. The idea behind a
balance plot is that means measure the weighted center, or a balance point, for
a set of values. Of course, if you have a plank with many weights on it, where
the balance point lies depends not only on the weights, but also on their
relative locations... the farther towards the ends of the plank a weight lies,
the more leverage it has. When describing diversity, there are simple rules for
the species weights ([relative] abundances) and where they lie (their *scaled*
rarities). The rarity scaling is given by the exponent $\ell$ from equation 1,
the argument `l`. Here, we create three balance plots for the three Pythagorean means, given when $\ell \in \{-1,0,1\}$: 

```{r Balance plot setup}

rich_bal <- rarity_plot(ab = uneven_comm, l = 1, base_size = 12) +
  ggplot2::labs(x = "") +
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2") 

shan_bal <- white_y(rarity_plot(ab = uneven_comm, l = 0, base_size = 12, verbose = F)) + #do not return all the text this time
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2")

simp_bal <- white_y(rarity_plot(ab = uneven_comm, l = -1, base_size = 12, verbose = F)) + 
  ggplot2::labs(x = "") +
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2")
```
In a balance plot, the species abundances are plotted on the y-axis, species
rarities on the x-axis (values are back-transformed, but scaling is dictated by
`l`), the mean is shown with a triangular "fulcrum," and reference points are
illustrated for the three Pythagorean means (i.e., arithmetic, geometric, and
harmonic, given when `l= c(1, 0, -1)`, and equal to richness, Hill-Shannon
diversity, and Hill-Simpson diversity). By default, each individual is plotted
as a "block", and species with the same rarity are stacked vertically.

```{r Balance plot print, fig.width = 3, fig.height = 3}
rich_bal 
shan_bal
simp_bal

```

## Simulating communities with known diversity

Unfortunately, it is difficult to assess the performance of diversity estimators
with real-world data, because ecological communities are not typically fully
censused, and therefore the true diversity of those communities is not known.
There are, of course, a few exceptional datasets... mostly observations of
vascular plants, often trees, where the species identity and abundances in each
plot is known (CITE tropical tree dataset). However, these may not be
representative of the variability in species abundance distributions found in
nature. To enable a more comprehensive view, `MeanRarity` includes a tool to
simulate species abundance distributions where diversity values are given as
parameters. This way, diversity (in this case, richness and Hill-Simpson
diversity) are known and fixed, rather than epiphenomena that fall out of the
simulations. In order to achieve this, we simulate species abundance
distributions with known, finite, true diversity, but infinite
abundance, from which finite samples may be taken. The diversity of a given
sample from such as SAD is, of course, not precisely pre-determined.

This approach differs from species abundance distribution simulations
implemented in other R packages. For example, `mobsim` [@May2018a] and `sads`
(CITE) allow users to specify continuous parametric probability distributions,
but then explicitly take finite, discrete samples from those distributions in
the simulation (in `mobsim` these can even be spatially explicit!). Users input
cannonical parameters of the probability distributions and the number of
individuals to sample. These are powerful simulation tools, but one drawback is
that the diversity resulting from any given set of parameter values is variable.

`fit_SAD()`, by contrast, does not simulate individuals at all. Instead, it
always gives precisely the same relative abundances for a given set of
parameters, which may be a helpful tool for exploring the behavior of diversity
metrics under different community size and sampling scenarios. 

`fit_SAD()` simulates a species abundance distribution with known richness,
Hill-Simpson diversity, and a distributional assumption (lognormal or gamma). It
returns a shape parameter for the chosen probability distribution (because we're
focused on relative abundances, multiplicative rescaling is irrelevant), the
richness, Hill-Shannon, and Hill-Simpson diversity of the simulated species
abundance distribution, and relative abundances for each simulated species. It
provides a warning when the distribution includes species with very low relative abundances, which implies a large total community abundance.

```{r fit species abundance distributions}
# set target richness and Hill-Simpson diversity levels
rich <- 20
simpson <- 12
even_SAD <- fit_SAD(rich = rich, simpson = simpson, distr = "lnorm")

#output is a list with three elements
even_SAD

rich <- 20
simpson <- 6
uneven_SAD <- fit_SAD(rich = rich, simpson = simpson, distr = "lnorm")



ev<-radplot(even_SAD$rel_abundances) + 
  ggplot2::labs(y = "relative abundance", title =NULL)+
  ggplot2::ylim(c(0, 0.4))
un<-radplot(uneven_SAD$rel_abundances) + 
  ggplot2::labs(ylab = "relative abundance", title =NULL) +
  ggplot2::ylim(c(0, 0.4))

ev
un
```

`fit_SAD()` generates an ideal species abundance distribution, with relative
abundances for a community with infinite abundance. To take a finite sample, we
use `sample_infinite()` (so named because it samples an infinite SAD;
`sample_finite()` is for sub-sampling from a realized community/finite sample).
The realized samples have Poisson-type noise about the mean species relative
abundances given by `fit_SAD()` (that is, they are samples from a multinomial
distribution). The diversity of a sample asymptotically approaches the simulated
diversity as sample size approaches infinity.

```{r sample from SAD}

# diversity of a small sample unlikely to match the simulated diversity values

# first, sample from our simulated SADs

even_sample <- sample_infinite(even_SAD$rel_abundances, 50)
uneven_sample <- sample_infinite(uneven_SAD$rel_abundances, 50)
#sample_infinite returns a numeric vector of sample species abundances
even_sample
uneven_sample

# compare diversity from sample vs. ideal SAD: 
# likely pretty far off for a small sample!

# with an even SAD, all diversities lower
even_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(even_sample, l)
}))

# with an uneven SAD, sample Hill-Simpson can be higher than true Hill-Simpson 
# richness is always lower
uneven_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(uneven_sample, l)
}))


big_even_sample <- sample_infinite(even_SAD$rel_abundances, 1e7)
big_uneven_sample <- sample_infinite(uneven_SAD$rel_abundances, 1e7)

# with a big enough sample, simulated diversities nearly recovered
even_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(big_even_sample, l)
}))

# with an uneven SAD, sample Hill-Simpson can be higher than true Hill-Simpson 
# richness is always lower
uneven_SAD$community_info
rbind(lapply(c(1, 0, -1), function(l){
  rarity(big_uneven_sample, l)
}))

```

For any distribution of relative abundances, one can ask, "what is the Hill diversity of this distribution?" Yet the answer, of course, depends on the chosen value of the scaling parameter $\ell$. For a comprehensive answer, one might look across values of $\ell$, also known as a Hill "diversity profile" [@Chao2014b]. `MeanRarity` provides a function to generate a Hill diversity profile based on $\ell$ (rather than $q$): `divpro()`. 

```{r diversity profile}
# the even and uneven communities have same richness but different Hill diversity when `ell != 1`
ggplot2::ggplot(divpro(even_SAD[[3]]),
                ggplot2::aes(ell, d)) +
  ggplot2::geom_line() +
  ggplot2::ylim(0,22) +
  ggplot2::theme_classic() +
  ggplot2::labs(x = "scaling exponent \'l\'", y = "Hill diversity")

ggplot2::ggplot(divpro(uneven_SAD[[3]]),
                ggplot2::aes(ell, d)) +
  ggplot2::geom_line() +
    ggplot2::ylim(0,22) +
  ggplot2::theme_classic() +
  ggplot2::labs(x = "scaling exponent \'l\'", y = "Hill diversity")



```

# EXAMPLE APPLICATIONS

# ALTERNATIVE SOFTWARE


# CONCLUSIONS


## Acknowledgements

## Author contributions

M.R. and J.D. conceived the package concept and structure and generated source
code for package functions. M.R. implemented the first package version and the
shiny online application. M.R. wrote the manuscript with help from J.D.

# DATA ACCESSIBILITY
This article does not use any empirical data. The package is available on Github
(SOME KIND OF URL OR DOI REF).


```{r table of functions}
table_dat<-data.frame(
  `Function name` = c("rarity", "rarity_plot", "radplot", "divpro", "sample_infinite", "sample_finite", "fit_SAD")
  , `General purpose` = c("Computation", "Visualization", "Visualization", "Visualization", "Sampling", "Sampling", "Simulation")
  , `Description` = c("Compute Hill diversity using the rarity parameterization")
)
```








